% !TEX root = main.tex
Today's complex engineered systems (you can replace complex engineered systems with CPS) are characterized by the interaction of a large number of heterogeneous components. Consequently, the models used to analyze these systems are equally complex and consist of heterogeneous sub-models relying on different modeling assumptions and based on principles from different scientific disciplines. It is not uncommon to encounter a patchwork of differential equations, difference equations, hybrid automata, lookup tables, custom switching logic, low-level legacy code, etc. To further compound the difficulty in analyzing these systems, different components of a complex engineered system are typically designed by different suppliers. Although a high-level specification for these components may be known, detailed models are not available for intellectual property reasons. We are thus faced with a tremendous gap between the existing analysis techniques that rely on closed-form models and the models available in industry. It is, therefore, not surprising the emphasis that industry places on simulation since despite the complexity of models, it is always possible to simulate them. Therefore, it is a natural question to ask whether we can provide formal analyses about certain properties of these complex systems based solely on the information obtained via their simulations. In this paper, we focus on one of the most important of such properties in the context of control theory: stability.

More formally, we consider a dynamical system as in:
\begin{equation}\label{eq:dynamicalsystemGeneral}x_{k+1} = f(k, x_k),
\end{equation}
where, $x_k \in X$, $k \in \N$ is used to index time.
We start with the following question to serve as a stepping stone: Given N pairs, $(x_1, y_1)$, $(x_2, y_2)$, $\ldots$, $(x_N, y_N)$ such that $y_{k} = f(k, x_k)$, what can we say about the stability of the system \eqref{eq:dynamicalsystemGeneral}? For the rest of the paper, we use the term \emph{black-box} to refer to models where we do not have access to its dynamics, yet we can observe $f(x)$ by exciting it with different initial conditions $x$. \textcolor{red}{We should say that this is non-ideal when we don't know what the state is, but it is a start, and it makes sense in certain situations.} One approach to this problem is firstly identifying the dynamics, i.e., $f$ and then applying the existing techniques in the model-based stability analysis literature. However, unless $f$ is a linear function, there are two main reasons behind our quest to directly work on input-output pairs and bypassing the identification phase: (1) Even when the function $f$ is known, in general, the stability analysis is a very difficult problem \cite{stabilityHard1}, \cite{stabilityHard2}. (2) \textcolor{red}{Paulo wants to change this: The existing identification techniques can only identify $f$ up to an approximation error. How to relate this identification error to an error in the stability of the system \eqref{eq:dynamicalsystemGeneral} is still a nontrivial problem.}

The initial idea behind this paper was influenced by the recent efforts in \cite{topcu}, \cite{kapinski} and \cite{lazar} in using simulation traces to find Lyapunov functions for systems with known dynamics. \textcolor{red}{Will put Liberzon here.} In these works, the main idea is that if one can construct a Lyapunov function candidate decreasing along many finite trajectories starting from different initial conditions, it should also decrease along every other trajectory. Then, once a Lyapunov function candidate is constructed, this intuition is put to test by verifying the candidate function either via off-the-shelf tools as in \cite{topcu} and \cite{kapinski}, or via sampling based techniques as in \cite{lazar}. Note that, since we do not have access to the dynamics, the second step cannot be directly applied to black-box systems. However, these sampling based ideas raise the following question that we address in this paper: By observing that a candidate Lyapunov function decreases on a large number of simulations we empirically build a certain confidence that such candidate Lyapunov function is a bona-fide Lyapunov function. \emph{Can we translate this confidence into a confidence in the stability of the underlying system?} 

Note that, even in the case of a 2D linear system the connection between these two beliefs is nontrivial. In fact, one can easily construct an example with one stable and one unstable eigenvalue for which even though almost all trajectories diverge to the infinity, it is possible to construct a Lyapunov function candidate whose level sets are contracting everywhere except a small set. \textcolor{red}{Should we give a specific example here, and put a figure?} Moreover, the size of this "violating set" can be arbitrarily small based on the magnitude of the unstable eigenvalue.

In this paper, we take the first step to close this gap. Since the identification and stability analysis of linear systems are well understood, we do so by focusing on switched linear systems. Note that identification and deciding the stability of arbitrary switched linear systems is NP-hard \cite{jungersBook}. Aside from their theoretical value, switched systems model the behavior of dynamical systems in the presence of known or unknown varying parameters. These parameters can model internal properties of the dynamical system such as uncertainties, look-up tables, values in a discrete register as well as exogenous inputs provided by a controller in a closed-loop control system. \textcolor{red}{Need to make these examples more specific.}

The stability of switched systems closely is closely related to the \emph{joint spectral radius} (JSR) of the matrices appearing in \eqref{eq:switchedSystem}. Under certain conditions deciding stability amounts to deciding whether JSR is less than one \cite{jungersBook}. In this paper, we present an algorithm to bound the JSR of a switched linear system from $N$ observations. This algorithm is based on tools from the random convex optimization literature \cite{campi}, and provides an upper bound on the JSR with a user-defined confidence level. As $N$ increases, this bound gets tighter. Moreover, with a closed form expression, we characterize what the exact trade-off between the tightness of this bound and the number of samples is. In order to understand the quality of our upper bound, the algorithm also provides a deterministic lower bound.

The organization of the paper is as follows: \textcolor{red}{TO BE FILLED.}
%. Topcu et.al. \cite{topcu}, Kapinski et.al. \cite{kapinski}construct Lyapunov function candidates using the simulation traces, however to be able to formally verify the constructed Lyapunov function, they require the knowledge of the full dynamics. In \cite{lazar} and \cite{lazar2} Bobiti and Lazar address this and provide sampling based probabilistic and deterministic guarantees of a given Lyapunov function candidate.

% The answer is immediate when \eqref{eq:dynamicalsystemGeneral} is a linear time-invariant system because we can simply identify $f$ by $n$ linearly independent output trajectories.  



%In this paper, we seek the answer to this question for switched linear systems for which the problem immediately becomes nontrivial. A switched linear systems is in the form:
%\begin{equation}\label{eq:switchedSystem}x_{k+1} = A_{\sigma(k)}x_k,\end{equation}
%where, $\sigma: \N \to \{1,2, \ldots, m\}$ is the switching sequence and $A_{\sigma(k)} \in \calM$, for all $\sigma$ and $k$. Aside from their theoretical value, switched systems model the behavior of dynamical systems in the presence of known or unknown varying parameters. These parameters can model internal properties of the dynamical system such as uncertainties, look-up tables, values in a discrete register as well as exogenous inputs provided by a controller in a closed-loop control system. 

%\textcolor{red}{There is no model.
%System are getting more complex, multimodal, lookup tables, thermodynamics, there is a need to ?prove properties about these systems, despite the fact that the modesl are mo complex are partially knowni Subcomponetns come from  different companies and black box.
%We still need to prove properties.
%The most important property in terms of control: stability.
%Even in the case when we have a model the stability is hard, for example Raphael, therefore we shouldn't expect we are gonna do better. The first step towards just based on switched systems. Because these are the very simple class of systems, resoanable it captures many applications.
%Even in the case of linear systems we cannot say anything. Here, we put the example where even the system is linear, we cannot say anything by just looking at the lyapunov functions decrease.}


%To make our reasoning clearer, we introduce the \emph{Lyapunov exponent} of the system \eqref{eq:switchedSystem}, which is a numerical quantity describing its stability.
%\begin{definition} Given a dynamical system as in \eqref{eq:dynamicalsystem} its \emph{Lyapunov exponent} is given by
%$$\rho =\inf{\{r:\,\forall x_0, \exists C\in \re^+: \quad x(0)=x_0 \Rightarrow x(t)\leq Cr^t\}}. $$
%\end{definition}
%Under certain conditions, deciding stability amounts to decide whether $\rho<1.$  In order to understand the quality of our techniques, we will actually try to prove lower and upper bounds on $\rho.$ 
%Assessing the stability of nonlinear systems by leveraging simulations has been an active area of research in the recent years. Simulation data has been used in both construction and verification of Lyapunov functions. Topcu et.al. \cite{topcu} and Kapinski et.al. \cite{kapinski} construct Lyapunov function candidates using the simulation traces, however to be able to formally verify the constructed Lyapunov function, they require the knowledge of the full dynamics. In \cite{lazar} and \cite{lazar2} Bobiti and Lazar address this and provide sampling based probabilistic and deterministic guarantees of a given Lyapunov function candidate. The presented method requires the knowledge of how fast the output of the system can change as the initial condition changes, and moreover the number of required samples increases exponentially in the dimension of the state, $n$.

%The stability of switched systems closely relates closely to the \emph{joint spectral radius} (JSR) of the matrices appearing in \eqref{eq:switchedSystem}. Under certain conditions deciding stability amounts to deciding whether JSR less than one or not. There has been a lot of work on developing algorithms to approximate this quantity, when the matrices appearing in \eqref{eq:switchedSystem} are known. Therefore, our work is also connected to the identification of switched systems, since once the system \eqref{eq:switchedSystem} is identified one can then apply these well-established results. However, there are two main reasons behind our quest to directly work on input-output pairs and bypassing the identification phase: (1) Even when $\calM$ is known, approximating the JSR is NP-hard \cite{jungers}. (2) Identifying the set $\calM$ is also NP-hard. Therefore, the existing identification techniques can identify $\calM$ up to an approximation error. As a result, how to relate this identification error to an error on the stability of \eqref{eq:switchedSystem} is still nontrivial.

%In this paper, we present an algorithm to approximate the JSR of a switched linear system from $N$ input-output pairs. This algorithm provides an upper bound on the JSR with a user-defined confidence level. As the number of samples increases, this bound gets tight. Moreover, we characterize with a closed form expression what the exact trade-off between the tightness of this bound and the number of samples is. In order to understand the quality of our technique, the algorithm also provides a deterministic lower-bound.

%The organization of the paper is follows: Section \label{preliminaries} introduces our notation and definitions from the switched linear system literature to present our results, Section \label{problemDefinition} formalizes the problem definition, Section \label{upperBound} provides an algorithm to compute a deterministic lower bound and 
%
%In this paper, we consider discrete-time switched linear systems of the form:
%\begin{equation}\label{eq:dynamicalsystem}x_{k+1} = A_{\sigma(k)}x_k,
%\end{equation}
%where, $x_k \in \R^n$, k is index of time and $\sigma: \N \to \{1,2, \ldots, m\}$ is the switching sequence. Let $y_k := x_{k+1}$. 
%We ask the following question:
%\begin{centering}
%\emph{Given N input-output-matrix pairs, $(x_1, y_1)$, $(x_2, y_2)$, $\ldots$, $(x_N, y_N)$ such that
%\begin{equation}\label{eq:triples}y_{k} = A_{\sigma(k)}x_k,
%\end{equation} for some $\sigma(k),$ what can we say about the stability of the system \eqref{eq:dynamicalsystem}?
%\end{centering}
%The only instance where answer to this question has been clearly shown is for linear time-invariant systems.
%
%Note that if \eqref{eq:dynamicalsystem} is a linear time-invariant system with only one mode, this question is easily answered by observing $n$ linearly independent data points. Switched systems can be used to model the behavior of a system of interest for different values of a parameter that varies. This parameter can represent internal parameters such as model uncertainties, as well as exogenous parameters such as inputs provided by a controller in a closed-loop control system. 


%- Stability analysis of dynamical systems is a challenging task. \\
%- Most methods in literature rely on knowing the model. Cite \\
%- How to make the case about the assumption of switched linear yet, we don't construct a model? Here cite
%something that says that this is a hard problem in general. (What is the relationship between switching linear regression)
%- Here we present stability analysis from data, this is easy for linear systems. We present something for switch linear.



%I can refer to the condition number paper?